---
title: "HW 02 report "
author: "Arunangshu Sarkar"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
# Use rmarkdown::render("Code/Sarkar_Project_01.Rmd", output_file = "../Reports/Sarkar_Project_01.pdf") to knit the markdown file
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = "simple")
library(utils)
library(ggplot2)
library(tidyverse)
```

# Exercise 2

## Exercise 2A

Below is a graph showing IQ levels by location status:

```{r}
datatab <- read.csv("../DataRaw/lead-iq-01.csv")
ggplot(datatab, aes(Smelter, IQ)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("IQ levels by location")
```

## Exercise 2B

Below is a formatted table using the `kable()` function
```{r}
knitr::kable(datatab[c(1:8, 68:76), ], 
             caption = "A few rows for each location status",
             booktabs = TRUE)
```


## Exercise 2C

*Boxplots:* From the box plots, there doesn't seem to be much difference in IQ levels for each location status, though it's hard to interpret the plots due to one extreme outlier present in the "Far" group. Now, the IQ of a person cannot be near 1000, so there seems to be an error when the data was entered. 

*Table:* This nicely formatted table shows a few rows in the data set for each location status.

## Exercise 2D

Calculating the means by location status:

```{r}
mean_df <- datatab %>%
  group_by(Smelter) %>%
  summarise(mean = mean(IQ))
```

The mean IQ for the subjects who were near the smelter was `r round(mean_df$mean[mean_df$Smelter == "Near"], digits = 2)`, and for those far from the smelter was `r round(mean_df$mean[mean_df$Smelter == "Far"], digits = 2)`.

## Exercise 3

We notice that the maximum value of IQ in the data set is `r max(datatab$IQ)`, which is not possible. Actually, this entry should be 99.

Below, we change it to 99:

```{r}
for (i in 1:nrow(datatab)) {
  if (datatab$IQ[i] == 999) {
    datatab$IQ[i] <- 99
  }
}
```

Now, we can check the boxplots again to understand the data more:

```{r}
ggplot(datatab, aes(Smelter, IQ)) +
  geom_boxplot() +
  theme_bw() +
  ggtitle("IQ levels by location, updated to remove 999 outlier")
```

Now, we calculate the means again:

```{r}
mean_df <- datatab %>%
  group_by(Smelter) %>%
  summarise(mean = mean(IQ))
```

When the error is rectified, the mean IQ for the subjects who were near the smelter was `r round(mean_df$mean[mean_df$Smelter == "Near"], digits = 2)`, and for those far from the smelter was `r round(mean_df$mean[mean_df$Smelter == "Far"], digits = 2)`.

We now write the updated and processed data to the folder DataProcessed for later use:

```{r}
write.csv(datatab, "../DataProcessed/lead-iq-999removed.csv")
```

